A Linear Inside-Outside Algorithm for Correcting Sequencing Errors in Structured RNAs

### 2013-01-16 Recomb/methods_RECOMB.tex
Page limit

\subsection{Probabilistic Model}
Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
\begin{align*}
  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
\end{align*}
where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.

\subsubsection{Energy Contribution.}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}. Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. neither $\Gb\Ub$, $\Ub\Gb$, $\Cb\Gb$, $\Gb\Cb$, $\Ab\Ub$ nor $\Ub\Ab$). This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
\subsubsection{Isostericity Contribution.}
The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the contribution of the isostericity to the pseudo-energy as
\begin{align*}
  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
	\frac{
		\sum_{s'\in\Omega}
			\text{\ISO}((s_i',s_j'),(a,b))}
%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
{		
		|\Omega|
	}
\end{align*}
is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Isostericity scores range between $0$ and $9.7$, $0$ corresponding to a perfect isostericity, and a penalty of $10$ is used for missing entries.
The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.

@@

Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
%  under the SCFG derivation $S$ 
(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
We define a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
 desired probability,  the two  functions
$\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.

@@

Let us then remind the Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
\begin{equation}
  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
\end{equation}

\subsubsection{Inside computation.}
\begin{figure}[t]
\begin{minipage}{0.60\textwidth}
\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
\end{minipage}
\begin{minipage}{0.38\textwidth}
\caption{Principle of the inside computation (aka partition function). Any sequence with $m$ mutations over an interval $[i,j]$ 
can be decomposed as a sequence preceded by a, possibly mutated, base at $i$ 
(Unpaired case), a sequence surrounded by some base-pair (Stacking-pair case), 
or as two sequences separated by some base-pair (General base-pairing case). 
In each case, one has to investigate any possible ways to distribute mutations between the different sequences 
and locally instantiated bases.}
\end{minipage}
\end{figure}

The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e. the sum of Boltzmann factors, over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively). 
Such terms can be defined by recurrence, for which the following initial conditions holds:

\begin{equation}
	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
	\begin{array}{ll}
		1 &\text{If } m = 0\\
		0 &\text{Otherwise.}
	\end{array}\right.
\label{eq:Z_in}
\end{equation}
In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$. Since the energetic terms only depend on base pairs, they are not involved in the initial conditions.

The main recursion is composed of four terms:\\

\begin{equation}
\label{eq:Z_rec}
\resizebox{\textwidth}{!}{%
	$\Z{i,j}{m}{a,b}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} & \text{If }S_{i}=-1\\
      \displaystyle
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
			 \displaystyle
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
      \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
   		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
      \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
      \cdot\Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
      0 &\text{Otherwise}
	\end{array}\right.$}
\end{equation}
\
The cases can be broken down as follows:
\begin{description}
\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
any sequence consists in a, possibly mutated, nucleotide $a'$ at position  $i$, 
followed by a sequence over $[i+1,j]$ having either $m-\Kron_{a',s_i}$, accounting for a possible mutation 
at position $i$, and having flanking nucleotides $a'$ and $b$.
\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$ (to avoid exceeding the targeted distance $m$).
Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
then a stacking energy contribution is added. 
\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$. Since the number of mutations sum to $m$ over the whole sequence must , then a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
\end{description}

@@

\begin{minipage}{0.75\textwidth}

@@

\begin{minipage}{0.24\textwidth}

@@

@@

We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:

@@

	  \Z{j,n-1}{m}{X,X}

@@

The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
it is not base paired. If it is with a position $k>j$, we increase $j$ to include it.
 Thus, when we need
to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
structure between $(j,n-1)$ must be a set of independent stems. Therefore,
 all the outside energy between $[j,n-1]$ is
equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself unfolds as follows.

@@

only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{inside}

@@

\item[$-1<S_{i}<i$:]As above but if position $i$ is to pairing with a lower value.
\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 

@@

\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities.}
By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
be described in function of the \emph{inside} term as $\Z{0,n-1}{m}{X,X}$,
 for any nucleotide $X\in B$ or
in function of the \emph{outside} term, for any unpaired position $k$:
$$
	\Z{0,n-1}{m}{X,X}
	\equiv
	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a}
$$

We are now left to compute the probability that a given position is a given nucleotide.
We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
\begin{equation}
	\mathbb{P}(s_i = x\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [x]}}}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}
\end{equation}
where $\mathcal{W}{\substack{\ast\\ \ast}}$ is defined by:
\begin{equation}
\resizebox{\textwidth}{!}{%
$ \mathcal{W}^M_{\substack{i, [x]}} =  \left\{
	\begin{array}{ll}
			\sum_{m=0}^{M}
			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
		&\text{If }S_i = -1\\
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
			\cdot\Z{i+1,k-1}{m'}{x,b}
		&\text{If }S_i=k>i\\
    \displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
			\cdot\Z{k+1,i-1}{m'}{b,x}
		&\text{If }S_i=k<i
	\end{array}\right.
$}\label{eq:combine}
\end{equation}

>>>>

\subsection{Probabilistic Model}
Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
then any sequence $s$ has probability proportional to its Boltzmann factor
\begin{align*}
  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
\end{align*}
where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.

\subsubsection{Energy Contribution.}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, taking values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}. Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. neither $\Gb\Ub$, $\Ub\Gb$, $\Cb\Gb$, $\Gb\Cb$, $\Ab\Ub$ nor $\Ub\Ab$). This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed 
(e.g. multiloops do not account for base-specific contributions). Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
\subsubsection{Isostericity Contribution.}
The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the isostericity contribution to the pseudo-energy as
\begin{align*}
  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
	\frac{
		\sum_{s'\in\Omega}
			\text{\ISO}((s_i',s_j'),(a,b))}
%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
{		
		|\Omega|
	}
\end{align*}
is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Isostericity scores range between $0$ and $9.7$, $0$ corresponding to a perfect isostericity, and a penalty of $10$ is used for missing entries.
The isostericity contribution will favor exponentially sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.

@@

Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $m$ mutations from $s$
%  under the SCFG derivation $S$ 
(formally $\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
We define a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to compute these
 probability. 
%from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.

@@

Let us then remind the Hamming distance function $\delta: B^*\times B^* \to \mathbb{N}^+$, which takes two sequences $s'$ and $s''$ as input, $|s'|=|s''|$, and returns the number of differing positions.
Finally, let us denote by $E_{(i,j),ab \to ab'}^{\Omega,\beta}$ the local contribution of a base-pair $(i,j)$ to the pseudo-energy, such that
\begin{equation}
  E_{(i,j),ab \to a'b'}^{\Omega,\beta}  = \alpha \cdot\ES^\beta_{a b \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
\end{equation}
\subsubsection{Inside computation.}
\begin{figure}[t]
\begin{minipage}{0.60\textwidth}
\resizebox{\textwidth}{!}{\input{FigDPInside.pgf}}
\end{minipage}
\begin{minipage}{0.38\textwidth}
\caption{Principle of the inside computation (partition function). Any sequence with mutations  
can be decomposed as a sequence preceded by a, possibly mutated, base 
(Unpaired case), a sequence surrounded by some base-pair (Stacking-pair case), 
or as two sequences segregated by some base-pair (General base-pairing case), and mutations must be distributed between sub-sequences 
and instantiated bases.}
\end{minipage}
\SpaceCheating
\end{figure}

The \emph{Inside} function $\Z{i,j}{m}{a,b}$ is the partition function, i.e. the sum of Boltzmann factors over all sequences in the interval $[i,j]$, at distance $m$ of $s_{[i,j]}$, and having flanking nucleotides $a$ and $b$ (at positions $i-1$ and $j+1$ respectively). 
Such terms can be defined by recurrence, for which the following initial conditions holds:

\begin{equation}
	\forall i \in [0,n-1]:\, \Z{i+1,i}{m}{a,b}=\left\{
	\begin{array}{ll}
		1 &\text{If } m = 0\\
		0 &\text{Otherwise.}
	\end{array}\right.
\label{eq:Z_in}
\end{equation}
In other words, the set of sequences at distance $m$ of the empty sequence is either empty if $m>0$, or restricted to the empty sequence, having energy $0$, if $m=0$. Since the energetic terms only depend on base pairs, they are not involved in the initial conditions. 
The main recursion itself is composed of four terms:\\

\begin{equation}
\label{eq:Z_rec}
\resizebox{.93\textwidth}{!}{%
	$\Z{i,j}{m}{a,b}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}  
      \Z{i+1,j}{m-\Kron_{a',s_i}}{a',b} & \text{If }S_{i}=-1\\
      \displaystyle
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
			 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}
			 \cdot \Z{i+1,j-1}{m-\Kron_{a'b',s_is_j}}{a',b'}&
			 \text{Elif }S_i=j \land S_{i-1}=j+1\\
			 \displaystyle
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
      \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
   		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
      \cdot\Z{i+1,k-1}{m-\Kron_{a'b',s_is_k}-m'}{a',b'}
      \cdot\Z{k+1,j}{m'}{b',b} & \text{Elif }S_i=k \land i < k \leq j\\
      0 &\text{Otherwise}
	\end{array}\right.$}
\end{equation}
\
The cases can be broken down as follows:
\begin{description}
\item[$S_{i}=-1$:] If nucleotide at position $i$ is unpaired, then 
any sequence is a concatenation of a, possibly mutated, nucleotide $a'$ at position  $i$, 
followed by a sequence over $[i+1,j]$ having $m-\Kron_{a',s_i}$ mutations (accounting for a possible mutation 
at position $i$), and having flanking nucleotides $a'$ and $b$.
\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$ having distance $m-\Kron_{a'b',s_is_j}$.
Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
then a stacking energy contribution is added. 
\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$. Since the number of mutations sum to $m$ over the whole sequence must , then a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
\end{description}

@@

\begin{minipage}{0.79\textwidth}

@@

\begin{minipage}{0.20\textwidth}

@@

\SpaceCheating

@@

The resulting terms $\Y{i,j}{m}{a,b}$ can be computed by recurrence, using as initial conditions:

@@

	  \Z{j,n-1}{m}{X,X}.

@@

The recurrence below extends the interval $[i,j]$, by including $i-1$ when
position $i$ not base paired, or extended in both directions if $i$ is paired with a position $k>j$.
% Thus, when we need
%to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
%structure between $(j,n-1)$ must be a set of independent stems. Therefore,
% all the outside energy between $[j,n-1]$ is
%equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. 
The recursion itself unfolds as follows:

@@

only term contributing directly to the energy is the isostericity of base pair $(i,k)$. 
Therefore, we consider the outside interval $[i-1,k+1]$, multiplying it by the \emph{inside}

@@

\item[$-1<S_{i}<i$:]As above but if position $i$ is paired with a lower value.
\item[Else:] Any other derivation of the SCFG does not correspond to the 

@@

\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities.}
By construction, the partition function over all sequences at exactly $m$ mutations of a reference sequence $s$ can 
be either described from the \emph{inside} contribution $\Z{0,n-1}{m}{X,X}$ of the whole sequence,
$\forall X\in B$, or from \emph{outside} terms as:
$$
	\Z{0,n-1}{m}{X,X}
	\equiv
	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a},\forall k \text{	unpaired.}
$$

We are now left to compute the probability that a given position is a given nucleotide.
We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations, one defines
\begin{equation}
	\mathbb{P}(s_i = x\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [x]}}}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}
\end{equation}
where $\mathcal{W}{\substack{\ast\\ \ast}}$ is defined by:
\begin{equation}
\resizebox{\textwidth}{!}{%
$ \mathcal{W}^M_{\substack{i, [x]}} =  \left\{
	\begin{array}{ll}
			\sum_{m=0}^{M}
			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
		&\text{If }S_i = -1\\
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
			\cdot\Z{i+1,k-1}{m'}{x,b}
		&\text{If }S_i=k>i\\
    \displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
			\cdot\Z{k+1,i-1}{m'}{b,x}
		&\text{If }S_i=k<i
	\end{array}\right.
$}\label{eq:combine}
\end{equation}

### 2013-01-16 Recomb/results_RECOMB.tex
Page limit

\begin{table}[t]
\begin{tabular}{cp{27em}}
\begin{tabular}{lccc}
$|s|$&\multicolumn{3}{c}{\#mutations}\\\cline{2-4}
		 			  & 6   &  12  & 24\\\cline{2-4}
100  				& 35s  & 238s & 1023s\\
300  			& 135s & 594s &2460s\\\cline{2-4}
		 						& 25   &    &	50		\\\cline{2-4}
500         & 5400s&       &  21003s    \\
\end{tabular}&
\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
 of mutations. $\alpha$ is
set at $0.5$,  $\beta$ to $15$ and $|\Omega|=44$.}
\end{tabular}
\label{tab:time}
\end{table}

@@

To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
by \RNApyro. More specifically, we fixed a threshold $\lambda \in [0,1]$, and predicted an error at position $i$ in sequence $\omega$ if and only if the
probability $P(i,x)$ of a nucleotide $x \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides having probability
greated than $\lambda$, that is  $$C(i) = \{ x \; | \;  x \in \{ A,C,G,U \} \mbox{ and } P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$$ where $\omega[i]$ is
the nucleotide at position $i$ in the input sequence. We note that, for lower thresholds, multiple nucleotides may be available in $C(i)$ to correct
the sequence. Here, we remind that our aim is to estimate the potential of error-correction of \RNApyro, and not to develop a full-fledged error-correction pipe-line, which  
shall be the subject of further studies. Finally, we progressively varied $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
under the curve (AUC). Our results are reported in Figure~\ref{fig:ROCall}.

\begin{figure}
\begin{center}
	\includegraphics[width=\textwidth]{subfigs_perform.png}
\end{center}
%\begin{subfigure}[b]{0.3\textwidth}
%\includegraphics[width=1.2\textwidth]{figures/ROC_6.pdf}
%\caption{6 mutations}
%\label{fig:ROC6mut}
%\end{subfigure}
%\hfill
%\begin{subfigure}[b]{0.3\textwidth}
%\includegraphics[width=1.2\textwidth]{figures/ROC_12.pdf}
%\caption{12 mutations}
%\label{fig:ROC12mut}
%\end{subfigure}
%\hfill
%\begin{subfigure}[b]{0.3\textwidth}
%\includegraphics[width=1.2\textwidth]{figures/ROC_24.pdf}
%\caption{24 mutations}
%\label{fig:ROC24mut}
%\end{subfigure}
\caption{Performance of error-correction. Subfigures show accuracy with under-estimated error rates (6 mutations), exact estimates (12 mutations) and over estimates 
(24 mutations). We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values 
ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over the 10 experiments.}
\label{fig:ROCall}
\end{figure}

>>>>

\begin{table}[t]
\begin{minipage}{.4\textwidth}
\begin{tabular}{rccc}
Length &\multicolumn{3}{c}{\#mutations}\\\cline{2-4}
		 			  & 6   &  12  & 24\\\cline{2-4}
100  				& 35s  & 238s & 1023s\\
300  			& 135s & 594s &2460s\\\cline{2-4}
		 						& 25   &    &	50		\\\cline{2-4}
500         & 5400s&       &  21003s    \\
\end{tabular}
\end{minipage}
\begin{minipage}{.59\textwidth}
\caption{Time required by the computation of probabilities. First column indicates the length and  the column indexes indicate the number
 of mutations. $\alpha$ is
set at $0.5$,  $\beta$ to $15$ and $|\Omega|=44$.}
\label{tab:time}
\end{minipage}
\SpaceCheating
\end{table}

@@

To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
by \RNApyro. More specifically, we fixed a threshold $\lambda \in [0,1]$, and predicted an error at position $i$ in sequence $\omega$ if and only if the
probability $P(i,x)$ of a nucleotide $x \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides having probability
greated than $\lambda$, that is  $C(i) = \{ x \; | \;  x \in \{ A,C,G,U \} \mbox{ and } P(i,x) > \lambda \mbox{ and }  n \neq \omega[i] \},$ where $\omega[i]$ is
the nucleotide at position $i$ in the input sequence. We note that, for lower thresholds, multiple nucleotides may be available in $C(i)$ to correct
the sequence. Here, we remind that our aim is to estimate the potential of error-correction of \RNApyro, and not to develop a full-fledged error-correction pipe-line, which  
shall be the subject of further studies. Finally, we progressively varied $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
under the curve (AUC). Our results are reported in Figure~\ref{fig:ROCall}.

\begin{figure}
\centering
	\includegraphics[width=\textwidth]{subfigs_perform.png}\\

%\begin{subfigure}[b]{0.3\textwidth}
%\includegraphics[width=1.2\textwidth]{figures/ROC_6.pdf}
%\caption{6 mutations}
%\label{fig:ROC6mut}
%\end{subfigure}
%\hfill
%\begin{subfigure}[b]{0.3\textwidth}
%\includegraphics[width=1.2\textwidth]{figures/ROC_12.pdf}
%\caption{12 mutations}
%\label{fig:ROC12mut}
%\end{subfigure}
%\hfill
%\begin{subfigure}[b]{0.3\textwidth}
%\includegraphics[width=1.2\textwidth]{figures/ROC_24.pdf}
%\caption{24 mutations}
%\label{fig:ROC24mut}
%\end{subfigure}
\caption{Performance of error-correction. Subfigures show accuracy with under-estimated error rates (6 mutations), exact estimates (12 mutations) and over estimates 
(24 mutations). We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values 
ranging of $\alpha=\{0,0.5,0.8,1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over the 10 experiments.}
\label{fig:ROCall}\SpaceCheating
\end{figure}
